Malacology: A Programmable Storage System

### 2016-10-21 paper/paper.tex
paper: general shrinkage - squash deadline day commits - touch up 6.1 and fix captions - remove awkward duplicate active storage discussion - remove table I spent 4 months on - add citations for Lua, Paxos, chain replication - add carlos 2am feedback - consumed Domino's pizza with pineapple, pepperoni, cheese
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

@@

\title{Malacology: A Programmable Storage System}
\authorinfo{Paper ID: 165}
           {Total Page: 14}
\date{}
\begin{document}

@@

\begin{abstract}
Storage systems are caught between the need to evolve data processing applications efficiently and quickly,
and the increasing velocity with which storage device technology evolves. This puts tremendous
pressure on storage systems to support rapid change both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system. In this work, we introduce the concept of a programmable storage system, a system that exposes internal services and abstractions with the goal of re-using and building new services on top of the storage stack. We present Malacology, a prototype implementation of a programmable storage system based on Ceph
to explore how existing abstractions of common storage system services 
can be leveraged to 
adapt to the needs of
new data processing systems and the
increasing speed of storage devices. We show how this approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems. We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance
distributed shared-log that leverages flash devices. The evaluation
demonstrates that our services inherit desirable qualities of the back-end
storage system, including the ability to balance load, efficiently propagate
cluster metadata, recover from failure, and to navigate trade-offs between
latency and throughput using leases.
\end{abstract}

@@

The evolutionary pressure placed on storage systems by these trends raises the
question of whether there are principles that storage systems designers can
follow to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.  In this paper we
investigate an approach that focuses on identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse subsystems and leverage their
optimizations, established correctness, robustness, and efficiency.

{\it \textbf{We define a programmable storage system}} to be a storage system
that facilitates the re-use and extension of existing storage abstractions
provided by the underlying software stack, to enable the creation of new
services via composition.  A programmable storage system can be realized by
exposing existing functionality (such as metadata services and synchronization
and monitoring capabilities) as interfaces that can be ``glued together'' in a
variety of ways using a high-level language. In contrast to programmable
storage, \emph{active storage} differs through its focus on the injection and
execution of code within a storage system or storage device. Programmable
storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection and
execution of code within a storage system or storage device---in
that the
former is applicable to any component of the storage system, while the latter
focuses at the data access level (i.e. inject and execute arbitrary code at the
I/O access layers). Given this contrast, we can say that active storage is an
example of how one internal component (the storage layer) is exposed in a
programmable storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system that facilitates the
construction of new services by re-purposing existing subsystem abstractions of
the storage stack.  We build Malacology in Ceph, a popular open source software
storage stack.  We choose Ceph to demonstrate the concept of programmable
storage because it offers a broad spectrum of existing services, including
distributed locking and caching services provided by metadata servers,
durability and object interfaces provided by the backend object store, and
propagation of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. 
%Our contributions are:

% prototype programmable storage system that 
Malacology includes a non-exhaustive set of interfaces that can be used as
building blocks for constructing novel storage abstractions, including:

@@

\item An interface for \textbf{durability} that persists policies using the
underyling storage stack's object store.

@@

{\it \textbf{We implement two distributed services}} using Malacology to
demonstrate the feasibility of the programmable storage approach:

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
% and
%conclude.

@@

Building storage stacks from the ground up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like exposing data locality and relaxing
POSIX constraints to to achieve application-specific I/O
optimizations.~\cite{CITEME}. Alternatively, general-purpose storage stacks are
built with the flexiblity to serve many applications by providing a variety of
interfaces and tunable parameters. Unfortunately, managing competing forces in
these systems is difficult and users want more control from the general-purpose
storage stacks without going as far as building their storage system from the
ground up.

To demonstrate this trend in storage systems we examine the state of
programmability in Ceph. Something of a storage swiss army knife, Ceph
simultaneously supports file, block, and object interfaces on a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already provides
some degree of programmability; the OSDs support domain-specific object
interfaces that are implemented by composing existing low-level storage
interfaces that execute atomically. These interfaces are written in C++ and are
statically loaded into the system.

The Ceph community provides empirical evidence that developers are already
beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
shows a dramatic growth in the production use of domain-specific interfaces in
the Ceph community since 2010. In that figure, classes are functional groupings
of methods on storage objects (e.g. remotely computing and caching the checksum
of an object extent).  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that a
gap in existing interfaces are being addressed through ad-hoc approaches to
programmability. In fact, Table~\ref{table:objclasses} categorizes existing
interfaces and we clearly see a trend towards reusable services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
Since 2010, the growth in the number of co-designed object storage interfaces
in Ceph has been accelerating. This plot is the number of object classes (a
group of interfaces), and the total number of methods (the actual API
end-points).}
\label{fig:obj-int-dev-growth}
\end{figure}

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Example & \# \\ \hline
    Logging  & Geographically distribute replicas & 11 \\ \hdashline
    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
    Management & Scan extents for file system repair & \\ \hdashline
    Locking  & Grants clients exclusive access & 6 \\ \hdashline
    Other & Garbage collection, reference counting  & 4\\
\end{tabular}
\caption{A variety of object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
}
\label{table:objclasses}
\end{table}

The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the
custom object interface facility of Ceph could be due to a number of reasons,
such as the default algorithms/tunables of the storage system being
insufficient for the application's performance goals, programmers wanting to
exploit application-specific semantics, and/or programmers knowing how to
manage resources to improve performance. A solution based on
application-specific object interfaces is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

@@

\item Policies and optimizations are usually hard-wired into the services and
one has to be careful when factoring them out not to introduce additional bugs.
These policies and optimiations are usually cross-cutting solutions to concerns
or trade-offs that cannot be fully explored yet (as they relate to workload or
hardware). Given these policies and optimizations, decomposition of otherwise
orthogonal internal abstractions can be difficult or dangerous.

@@

Malacology implements and demonstrates the ideas of programmable storage in
Ceph. We choose Ceph because it is a production quality system and because it
is open source. The large developer community ensures that code is
robust and the visibility of the code lets us expose any interface we want. In
the next section we describe the Ceph components that we expose as Malacology
interfaces.

@@

\begin{table*}
\centering
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

The guiding principle
is to re-use existing services and extend them so that these services can be
\emph{programmed}. We accomplish programmability of a service by exporting
bindings (or ``hooks'') for an interpreted programming language so that
programming can occur without having to restart the storage system (see also
below,~\S\ref{sec:durability}).

There are multiple reasons that make Lua an attractive runtime for implementing
Malacology. Lua is a portable, embedded scripting language that offers superior
performance and productivity trade-offs, including a JIT-based implementation
that is well known for near native performance. Additionally, Lua has been used
extensively in game engines, and systems research \cite{neto:dls14-luaos},
including storage systems where it has been effectively used both on
\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
flexibility of the runtime allows execution sandboxing in order to address
security and performance concerns.

We will now discuss the common subsystems used to manage storage system and how
Malacology makes them programmable.

@@

% straw man example
Keeping track of the state of a distributed system is an essential part of any
successful service and a necessary component in order to diagnose and detect
failures, when they occur. This is further complicated by variable propagation
delays and heterogenous hardware in highly dynamic environments.

In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem data structures
called ``maps'' that record membership and status information.  A Paxos-based
monitoring service is responsible for integrating state changes into cluster
maps, responding to requests from out-of-date clients and synchronizing members
of the cluster whenever there is a change in a map so that they all observe the
same system state. As a fundamental building block of many system designs,
consensus abstractions such as Paxos are a common technique for maintaining
consistent data versions, and are a useful system to expose.

The default behavior of the monitor can be seen as a Paxos-based notification
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
clients to identify when new values (termed epochs in Ceph) are associated to
given maps.  While Ceph does not expose this service directly, a key-value
service designed for managing configuration metadata that is built on top of
the consensus engine is an example of a high-level service within Ceph that
re-uses existing abstractions.  While a key-value service is generally useful,
it doesn't provide many of the useful services hidden within the monitoring
framework that would be required for applications with more demanding
requirements, such as creating arbitrary maps with application-specific
semantics and interfaces. Since the monitor is intended to be out of
high-performance I/O paths, a general guideline is to make use of this
functionality infrequently and to assign small values to maps.

\paragraph*{Malacology:} Malacology exposes a strongly-consistent view of
time-varying service metadata as a service rather than a hidden internal
component. Malacology provides a generic API for adding arbitrary values to
existing subsystem cluster maps. As a consequence of this, applications can
define simple but useful service-specific logic to the strongly-consistent
interface, such as authorization control (just specific clients can write new
values) or to trigger actions based on specific values (e.g. sanitize values).
The higher level services we implement in \S\ref{sec:services} make use of this
functionality to register, version and propagate dynamic code (Lua scripts) for
new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
policies in the load balancer \S\ref{malacology:mds}).  Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughout the cluster so that
clients are properly synchronized with the latest interfaces.

@@

Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
supports the creation of application-specific object
interfaces~\cite{weil_rados_2007}. The ability to offload computation can
reduce data movement, and transactional interfaces can significantly simplify
construction of complex storage interfaces that require uncoordinated parallel
access.

An object interface is a plugin structured in a similar way to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and write to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
object interfaces may be transactionally composed along with application
specific logic to create interfaces with a rich set of semantics.  An example
is the use of a read-modify-write operation to perform consistent object
updates across the byte stream and key-value database interfaces without
relying on complex client-driven concurrency control.

\begin{figure}[htbp]
\centering
\includegraphics{figures/implementation.png}
\caption{Malacology re-uses Ceph subsystems to allow users to dynamically
define new object/data interfaces and load balancing policies. It achieves this
by extending the object storage daemon (OSD) and metadata server daemon (MDS)
subsystems with an embedded Lua VM.  It uses the of the strongly-consistent
service metadata interface to propogate the object interfaces and versions
across the cluster.
\label{fig:implementation}} \end{figure}

@@

To address these concerns, Malacology takes advantage of Lua extensions
contributed by the Ceph community. This allows new object interfaces to be
dynamically loaded into the system and modified at runtime, resulting in a
object storage API with economy of expression, which at the same time provides
the full set of features of the base object class. New object interfaces that
are expressed in thousands of lines of code can be implemented in approximately
an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
Lua does not prevent deployment of malicious code, certain types of coding
mistakes can handled gracefully, and access policies are used to restrict the
feature to trusted users.

@@

The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
file systems protect resources by providing hierarchical indexing and
distributed locking services.  In Ceph, the locking service implements a
capability-based system that expresses what data and metadata clients are
allowed to access as well as what state they may cache and modify locally.
While designed for a fixed file abstraction, indexing, locking, and caching are
all common services that are useful to a broad spectrum of applications.
Distributed applications that share centralized resources (e.g. a database or
directory) face similar challenges which are often solved using
application-specific sharding.

Ceph also addresses the challenge of balancing metadata load with a separate
metadata cluster. This cluster uses load balancing policies to migrate
directory inodes around the cluster to alleviate overload at single nodes. The
policies use metrics based on system state (e.g.  CPU and memory utilization)
and statistics collected by the cluster (e.g. the popularity of an inode). Ceph
uses dynamic subtree partitioning to move variable sized namespace
subtrees~\cite{weil:sc2004-dyn-metadata}. These units can be shipped anywhere
(i.e., to any metadata server of any capacity) at any time for any reason. The
original balancer was designed with hard-coded policies and tunables.

@@

\paragraph*{Malacology:} interfaces are added in strategic spots in the
metadata service for guarding shared resources, defining new file types,
and load balancing.

@@

{\bf File type interface.} The metadata service manages a POSIX file system
hierarchy where files and directories are represented as inode data structures
and expose a POSIX file interface. Application that manage large amounts of
metadata (e.g. users or database snapshots) often require a naming service. We
allow new inode types to be defined such that applications can create
domain-specific interfaces to inodes that may modify locking and capability
policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
distributed shared-log built on Malacology.

{\bf Load balancing interface.} Mantle~\cite{sevilla:sc15-mantle}, a
programmable metadata load balancer, has been implemented on Malacology so that
its infrastructure can enjoy all the properties of existing internal
abstractions.  The load balancing logic is intact but the infrastructure has
been improved to safely load, manage, and persist its load balancing policies.
Composing Mantle using the service metadata and durability interfaces, Mantle
can safely version balancer policies, save balancer policies in the backend
object store and centralize warnings/errors. When combined with file types,
load balancing policies can express a variety of rules such as multi-tentant
and multi-application balancing.

@@

Storing the balancers in RADOS is only possible because we use Lua as the
language for writing balancer code. If we used a language that needs to be
compiled, like the C++ object classes in the OSD, we would need to ensure
binary compatibility, which is complicated by different operating systems,
distributions, and compilers.

>>>>

\setlength{\belowcaptionskip}{-10pt}

@@

\title{Malacology: A Programmable Storage System}
\authorinfo{Paper ID: 165}
           {Total Pages: 12}
\date{}
\begin{document}

@@

\begin{abstract} Storage systems are caught between the need to evolve data
processing applications efficiently and quickly, and the increasing velocity
with which storage device technology evolves. This puts tremendous pressure on
storage systems to support rapid change both in terms of their interfaces and
their performance. But adapting storage systems can be difficult because
unprincipled changes might jeopardize years of code-hardening and performance
optimization efforts that were necessary for users to entrust their data to the
storage system. We introduce the concept of a programmable storage system, a
system that exposes internal services and abstractions with the goal of
re-using and building new services on top of the storage stack. We present
Malacology, a prototype implementation of a programmable storage system to
explore how existing abstractions of common storage system services can be
leveraged to adapt to the needs of new data processing systems and the
increasing variety of storage devices. We show how this approach allows a large
degree of flexibility for storage systems to evolve without sacrificing the
robustness of their code-hardened subsystems. We illustrate the advantages and
challenges of programmability by composing existing primitives into two new
higher-level services: a metadata load balancer and a high-performance
distributed shared-log. The evaluation demonstrates that our services inherit
desirable qualities of the back-end storage system, including the ability to
balance load, efficiently propagate cluster metadata, recover from failure, and
to navigate trade-offs between latency and throughput using leases.
\end{abstract}

@@

The evolutionary pressure placed on storage systems by these trends raises the
question of whether there are principles that storage systems designers can
follow to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.  In this paper we
investigate an approach that focuses on identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstractions one can reuse subsystems and leverage their
optimizations, established correctness, robustness, and efficiency.

{\it \textbf{Contribution 1:}} We define a programmable storage system to be a
storage system that facilitates the re-use and extension of existing storage
abstractions provided by the underlying software stack, to enable the creation
of new services via composition.  A programmable storage system can be realized
by exposing existing functionality (such as metadata services and
synchronization and monitoring capabilities) as interfaces that can be ``glued
together'' in a variety of ways using a high-level language. Programmable
storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection
and execution of code within a storage system or storage device---in that the
former is applicable to any component of the storage system, while the latter
focuses at the data access level. Given this contrast, we can say that active
storage is an example of how one internal component (the storage layer) is
exposed in a programmable storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system that facilitates the
construction of new services by re-purposing existing subsystem abstractions of
the storage stack.  We build Malacology in Ceph, a popular open source software
storage stack.  We choose Ceph to demonstrate the concept of programmable
storage because it offers a broad spectrum of existing services, including
distributed locking and caching services provided by metadata servers,
durability and object interfaces provided by the backend object store, and
propagation of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}).  Malacology is expressive enough to provide the
functionality necessary for implementing new services.

Malacology includes a non-exhaustive set of interfaces that can be used as
building blocks for constructing novel storage abstractions, including:

@@

\item An interface for \textbf{durability} that persists policies using the
underlying storage stack's object store.

@@

{\it \textbf{Contribution 2:}} We implement two distributed services using Malacology to
demonstrate the feasibility of the programmable storage approach:

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source software community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
% and
%conclude.

@@

Building storage stacks from the ground up for a specific purpose results in
the best performance. For example, GFS~\cite{ghemawat:sosp03} and
HDFS~\cite{shvachko:msst10} were designed specifically to serve Map/Reduce and
Hadoop jobs, and use techniques like exposing data locality and relaxing POSIX
constraints to achieve application-specific I/O optimizations. Another example
is Boxwood~\cite{maccormick:osdi04}, which experimented with B-trees
and chunk stores as storage abstractions to simplify application building.
Alternatively, general-purpose storage stacks are built with the flexibility to
serve many applications by providing a variety of interfaces and tunable
parameters.  Unfortunately, managing competing forces in these systems is
difficult and users want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

To demonstrate a recent trend towards more application-specific storage systems
we examine the state of programmability in Ceph. Something of a storage Swiss
army knife, Ceph simultaneously supports file, block, and object interfaces on
a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of object storage
devices (OSDs) that provide Ceph with data durability and integrity using
replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already
provides some degree of programmability; the OSDs support domain-specific
object interfaces that are implemented by composing existing low-level storage
interfaces that execute atomically. These interfaces are written in C++ and are
statically loaded into the system.

The Ceph community provides empirical evidence that developers are already
beginning to embrace programmable storage. Figure~\ref{fig:obj-int-dev-growth}
shows a dramatic growth in the production use of domain-specific interfaces in
the Ceph community since 2010. In that figure, classes are functional groupings
of methods on storage objects (e.g. remotely computing and caching the checksum
of an object extent).  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that
gaps in existing interfaces are being addressed through ad-hoc approaches to
programmability. In fact, Table~\ref{table:objclasses} categorizes existing
interfaces and we clearly see a trend towards reusable services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object storage interfaces
in Ceph has been accelerating. This plot is the number of object classes (a
group of interfaces), and the total number of methods (the actual API
end-points).}
\label{fig:obj-int-dev-growth}
\end{figure}

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Example & \# \\ \hline
    Logging  & Geographically distribute replicas & 11 \\ \hdashline
    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
    Management & Scan extents for file system repair & \\ \hdashline
    Locking  & Grants clients exclusive access & 6 \\ \hdashline
    Other & Garbage collection, reference counting  & 4\\
\end{tabular}
\caption{A variety of object storage classes exist to expose interfaces
    to applications. \# is the number of methods that implement these categories.
}
\label{table:objclasses}
\end{table}

The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the
custom object interface facility of Ceph could be due to a number of reasons,
such as the default algorithms/tunables of the storage system being
insufficient for the application's performance goals, programmers wanting to
exploit application-specific semantics, and/or programmers knowing how to
manage resources to improve performance. A solution based on
application-specific object interfaces is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
the application is CPU or IO bound, if it has locality, if its size has the potential to
overload a single node, etc.  Programmers often know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

@@

\item Policies and optimizations are usually hard-wired into the services and
one has to be careful when factoring them to avoid introducing additional bugs.
These policies and optimizations are usually cross-cutting solutions to
concerns or trade-offs that cannot be fully explored at the time the code is
written (as they relate to workload or hardware). Given these policies and
optimizations, decomposition of otherwise orthogonal internal abstractions can
be difficult or dangerous.

@@

The Malacology prototype we present has been implemented on Ceph.
While there are other systems on top of which Malacology could be implemented (see Table~\ref{table:examples}), we choose Ceph because it is a production quality system and because it
is open source. The large developer community ensures that code is
robust and the visibility of the code lets us expose any interface we want. In
the next section we describe the Ceph components that we expose as Malacology
interfaces.

@@

\begin{table*}
\centering\small
\begin{tabular}{  l | l | l | l }
%\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,burrows_chubby_2006}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{Common internal abstractions. The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces discussed in this paper.}
\label{table:examples}
\end{table*}

The guiding principle
is to re-use existing services and extend them so that these services can be
\emph{programmed}. We accomplish programmability of a service by exporting
bindings (or ``hooks'') for an interpreted programming language so that
programming can occur without having to restart the storage system (see also
below,~\S\ref{sec:durability}). Table~\ref{table:examples} shows the internal services from Ceph that we expose in the Malacology prototype via Lua~\cite{ierusalimschy1996lua} bindings.

Lua is a portable embedded scripting language and we choose it as the
interpreted language for Malacology because it offers superior performance and
productivity trade-offs, including a JIT-based implementation that is well
known for near native performance. Additionally, Lua has been used extensively
in game engines, and systems research \cite{neto:dls14-luaos}, including
storage systems where it has been effectively used both on
\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010}
and off \cite{sevilla:sc15-mantle} the performance critical path. Finally, the
flexibility of the runtime allows execution sandboxing in order to address
security and performance concerns. We will now discuss the common subsystems
used to manage storage system and how Malacology makes them programmable.

@@

% straw man example
Keeping track of state in a distributed system is an essential part of any
successful service and a necessary component in order to diagnose and detect
failures, when they occur. This is further complicated by variable propagation
delays and heterogeneous hardware in dynamic environments.

In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ({\it e.g.} the
metadata server) data structures called ``maps'' that record membership and
status information.  A Paxos~\cite{lamport_parttime_1998} monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as Paxos are a common technique for maintaining consistent
data versions, and are a useful system to expose.

The default behavior of the monitor can be seen as a Paxos-based notification
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
clients to identify when new values (termed epochs in Ceph) are associated to
given maps.  Since Ceph does not expose this service directly, as part of our Malacology implementation, we expose a key-value
service designed for managing configuration metadata that is built on top of
the consensus engine. Since the monitor is intended to be out of the
high-performance I/O path, a general guideline is to make use of this
functionality infrequently and to assign small values to maps.

\paragraph*{Malacology:} Malacology exposes a strongly-consistent view of
time-varying service metadata as a service rather than a hidden internal
component. Malacology provides a generic API for adding arbitrary values to
existing subsystem cluster maps. As a consequence of this, applications can
define simple but useful service-specific logic to the strongly-consistent
interface, such as authorization control (just specific clients can write new
values) or triggering actions based on specific values (e.g. sanitize values).
The higher level services we implement in \S\ref{sec:services} make use of this
functionality to register, version and propagate dynamic code (Lua scripts) for
new object interfaces defined in storage daemons (\S\ref{object-data-interface}) and
policies in the load balancer \S\ref{malacology:mds}).  Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughout the cluster so that
clients are properly synchronized with the latest interfaces.

@@

Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
supports application-specific object
interfaces~\cite{weil_rados_2007}. The ability to offload computation can
reduce data movement, and transactional interfaces can significantly simplify
construction of complex storage interfaces that require uncoordinated parallel
access.

An object interface is a plugin structured similarly to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output.  A wide range of native interfaces are available to
developers such as reading and writing to a byte stream, controlling object
snapshots and clones, and accessing a sorted key-value database. These native
interfaces may be transactionally composed along with application
specific logic to create semantically rich interfaces. An example would be
an interface that atomically updates a matrix stored in the bytestream and an
index of the matrix stored in the key-value database.

\begin{figure}[tbp]
\centering
\includegraphics{figures/implementation.png}
\caption{Malacology allows users to dynamically define object/metadata
interfaces by extending the object storage daemon (OSD) and metadata server
(MDS) subsystems with an embedded Lua VM.  It uses the service metadata
interface (M) to propagate interfaces/versions across the cluster
\label{fig:implementation}}
\end{figure}

@@

To address these concerns, Malacology takes advantage of Lua extensions
contributed by the Ceph community. This allows new object interfaces to be
dynamically loaded into the system and modified at runtime, resulting in a
object storage API with economy of expression, which at the same time provides
the full set of features of the base object class. New object interfaces that
are expressed in thousands of lines of code can be implemented in approximately
an order of magnitude less code~\cite{geambasu_comet_2010}. While the use of
Lua does not prevent deployment of malicious code, certain types of coding
mistakes can be handled gracefully, and access policies are used to limit access
to trusted users~\cite{ierusalimschy1996lua}.

@@

The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
file systems protect resources by providing hierarchical indexing and
distributed locking services.  In Ceph, the locking service implements a
capability-based system that expresses what data and metadata clients are
allowed to access as well as what state they may cache and modify locally.
While designed for the file abstraction, indexing, locking, and caching are
all common services that are useful to a broad spectrum of applications.
Distributed applications that share centralized resources (e.g. a database or
directory) face similar challenges which are often solved using
application-specific sharding.

Ceph also addresses the challenge of balancing metadata load with a separate
metadata cluster. This cluster uses load balancing policies to migrate
directory inodes around the cluster to alleviate overload at single
nodes~\cite{weil:sc2004-dyn-metadata}. The policies use metrics based on system
state (e.g.  CPU and memory utilization) and statistics collected by the
cluster (e.g. the popularity of an inode). Ceph uses dynamic subtree
partitioning to move variable sized namespace subtrees. These units can be
shipped anywhere (i.e., to any metadata server of any capacity) at any time for
any reason. The original balancer was designed with hard-coded policies and
tunables.

@@

\paragraph*{Malacology:} Interfaces are added in strategic spots in the
metadata service for guarding shared resources, defining new file types,
and load balancing.

@@

{\bf File type interface.} The metadata service manages a POSIX file system
hierarchy where files and directories are represented as inode data structures
and expose a POSIX file interface. Applications that manage large amounts of
metadata (e.g. users or database snapshots) often require a naming service. We
allow new inode types to be defined such that applications can create
domain-specific interfaces to inodes that may modify locking and capability
policies. We will show how this is used in Section~\ref{sec:seq} when we discuss a
distributed shared-log built on Malacology.

{\bf Load balancing interface.} Mantle~\cite{sevilla:sc15-mantle}, a
programmable metadata load balancer, has been implemented on Malacology so that
its infrastructure can enjoy all the properties of existing internal
abstractions.  The load balancing logic is intact but the infrastructure has
been improved to safely load, manage, and persist its load balancing policies.
Composing Mantle with the service metadata and durability interfaces, Mantle
can safely version balancer policies, save balancer policies in the backend
object store and centralize warnings/errors. When combined with file types,
load balancing policies can express a variety of rules such as multi-tenant
and multi-application balancing.

@@

Storing the balancers in RADOS is simplified by the use of an interpreted
language for writing balancer code. If we used a language that needs to be
compiled, like the C++ object classes in the OSD, we would need to ensure
binary compatibility, which is complicated by different operating systems,
distributions, and compilers.
